rm(list=ls())

Load packages
```{r}
library(readxl)
library(dplyr)
library(glmmTMB)
library(factoextra)
library(ggplot2)
library(piecewiseSEM)
library(ggplot2)
library(DHARMa)
library(emmeans)
library(gridExtra)
library(viridis)
library(gridExtra)
library(grid)
library(mgcv)
library(patchwork)
library(MuMIn)
library(ggeffects)
library(ggpubr)
```

Load data
```{r}
df <- read_excel("G:/My Drive/PhD/Data/Big data set/big_data_ch3_310325.xlsx", sheet = 2)
```

Make modifications to the data
```{r}
# throw out the lines for which there is no biomass data
df2 <- df %>% filter(!is.na(biomass))

# remove un-used columns
df2$airtemp_ave <- NULL
df2$airtemp_sd <- NULL
df2$soiltemp_ave <- NULL
df2$soiltemp_sd <- NULL
df2$comments <- NULL

#current dataset is 371 obs and 22 variables. Remove all NAs
df2 <- na.omit(df2)


### now split 2022 and 2023. I will work from df2 (347 obs.)
df2_2022 <- df2 %>% filter(year == 2022)
df2_2023 <- df2 %>% filter(year == 2023)
```

Make the seasons
```{r}
df2_2022 <- df2_2022 %>%
  mutate(season = case_when(
    weeknr < 23 ~ "early",
    weeknr >= 23 & weeknr <= 29 ~ "mid",
    weeknr > 29 ~ "late"
  ))

df2_2023 <- df2_2023 %>%
  mutate(season = case_when(
    weeknr < 23 ~ "early",
    weeknr >= 23 & weeknr <= 29 ~ "mid",
    weeknr > 29 ~ "late"
  ))

df2_2022$season <- as.factor(df2_2022$season)
df2_2023$season <- as.factor(df2_2023$season)
```

Make rainfall and windspeed into classes
```{r}
#rainfall
df2_2022$rainfall_cat <- ifelse(
  df2_2022$rainfall < 10, "Low",
  ifelse(df2_2022$rainfall <= 50, "Medium", "High")
)

df2_2023$rainfall_cat <- ifelse(
  df2_2023$rainfall < 10, "Low",
  ifelse(df2_2023$rainfall <= 50, "Medium", "High")
)


#windspeed
#windspeed * 0.1, because of the way it was noted in the KNMI dataset
df2_2022$wind_speed_ave <- df2_2022$wind_speed_ave * 0.1
df2_2023$wind_speed_ave <- df2_2023$wind_speed_ave * 0.1


#from m/s to km/h
df2_2022$wind_speed_ave_kmh <- df2_2022$wind_speed_ave * 3.6
df2_2023$wind_speed_ave_kmh <- df2_2023$wind_speed_ave * 3.6

#wind speed classes made according to :https://www.rmets.org/metmatters/beaufort-wind-scale


df2_2022$windspeed_cat <- ifelse(
  df2_2022$wind_speed_ave_kmh >= 12 & df2_2022$wind_speed_ave_kmh < 20, "Gentle breeze",
  ifelse(df2_2022$wind_speed_ave_kmh >= 20 & df2_2022$wind_speed_ave_kmh < 29, "Moderate breeze",
    ifelse(df2_2022$wind_speed_ave_kmh >= 29 & df2_2022$wind_speed_ave_kmh <= 38, "Fresh breeze",NA)
  )
)


df2_2023$windspeed_cat <- ifelse(
  df2_2023$wind_speed_ave_kmh >= 12 & df2_2023$wind_speed_ave_kmh < 20, "Gentle breeze",
  ifelse(df2_2023$wind_speed_ave_kmh >= 20 & df2_2023$wind_speed_ave_kmh < 29, "Moderate breeze",
    ifelse(df2_2023$wind_speed_ave_kmh >= 29 & df2_2023$wind_speed_ave_kmh <= 38, "Fresh breeze", NA)
  )
)
```

Scale all variables
```{r}
df2_2022$biomass.sc <- scale(df2_2022$biomass)
df2_2022$veghgt_ave.sc <- scale(df2_2022$veghgt_ave)
df2_2022$sm_ave.sc <- scale(df2_2022$sm_ave)
df2_2022$soilres_ave.sc <- scale(df2_2022$soilres_ave)
df2_2022$LUI2016.sc <- scale(df2_2022$LUI2016)
df2_2022$GPI200_ave.sc <- scale(df2_2022$GPI200_ave)
df2_2022$mowing.sc <- scale(df2_2022$mowing)
df2_2022$knmitemp_ave.sc <- scale(df2_2022$knmitemp_ave)

df2_2023$biomass.sc <- scale(df2_2023$biomass)
df2_2023$veghgt_ave.sc <- scale(df2_2023$veghgt_ave)
df2_2023$sm_ave.sc <- scale(df2_2023$sm_ave)
df2_2023$soilres_ave.sc <- scale(df2_2023$soilres_ave)
df2_2023$LUI2016.sc <- scale(df2_2023$LUI2016)
df2_2023$GPI200_ave.sc <- scale(df2_2023$GPI200_ave)
df2_2023$mowing.sc <- scale(df2_2023$mowing)
df2_2023$knmitemp_ave.sc <- scale(df2_2023$knmitemp_ave)
```

Make a dataset per season
```{r}
df2_2022_early <- subset(df2_2022, weeknr >= 14 & weeknr <= 22)
df2_2022_mid   <- subset(df2_2022, weeknr >= 23 & weeknr <= 29)
df2_2022_late  <- subset(df2_2022, weeknr >= 30 & weeknr <= 35)

df2_2023_early <- subset(df2_2023, weeknr >= 15 & weeknr <= 22)
df2_2023_mid   <- subset(df2_2023, weeknr >= 23 & weeknr <= 29)
df2_2023_late  <- subset(df2_2023, weeknr >= 30 & weeknr <= 34)
```






#Part III
Structural equation model
6 SEMs early, mid, late, 2022, 2023



Biomass 2022 - early 
```{r}
sem_biomass_early2022 <- glmmTMB(biomass ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_early, family = gaussian(), REML=FALSE)
summary(sem_biomass_early2022) #AIC=500.2
plot(simulateResiduals(sem_biomass_early2022, refit=FALSE))

sem_biomass_early2022_sqrt <- glmmTMB(sqrt(biomass) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_early, family = gaussian(), REML=FALSE)
summary(sem_biomass_early2022_sqrt) #AIC=231.2
plot(simulateResiduals(sem_biomass_early2022_sqrt, refit=FALSE))

sem_biomass_early2022_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_early, family = gaussian(), REML=FALSE)
summary(sem_biomass_early2022_log) #AIC=157.9
plot(simulateResiduals(sem_biomass_early2022_log, refit=FALSE))


AICc(sem_biomass_early2022, sem_biomass_early2022_sqrt, sem_biomass_early2022_log)

#best model
sem_biomass_early2022_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_early, family = gaussian(), REML=T)
summary(sem_biomass_early2022_log) 
plot(simulateResiduals(sem_biomass_early2022_log, refit=FALSE))
```

biomass 2022 - mid
```{r}
sem_biomass_mid2022 <- glmmTMB(biomass ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_mid, family = gaussian(), REML=FALSE)
summary(sem_biomass_mid2022) #AIC=413.1
plot(simulateResiduals(sem_biomass_mid2022, refit=FALSE))

sem_biomass_mid2022_sqrt <- glmmTMB(sqrt(biomass) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_mid, family = gaussian(), REML=FALSE)
summary(sem_biomass_mid2022_sqrt) #AIC=172.9
plot(simulateResiduals(sem_biomass_mid2022_sqrt, refit=FALSE))

sem_biomass_mid2022_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_mid, family = gaussian(), REML=FALSE)
summary(sem_biomass_mid2022_log) #AIC=76.5
plot(simulateResiduals(sem_biomass_mid2022_log, refit=FALSE))


AICc(sem_biomass_mid2022, sem_biomass_mid2022_sqrt, sem_biomass_mid2022_log)

#best model
sem_biomass_mid2022_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_mid, family = gaussian(), REML=T)
summary(sem_biomass_mid2022_log) #AIC=99
plot(simulateResiduals(sem_biomass_mid2022_log, refit=FALSE))
```

biomass 2022 - late
```{r}
sem_biomass_late2022 <- glmmTMB(biomass ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_late, family = gaussian(), REML=F)
summary(sem_biomass_late2022) #AIC=345.9
plot(simulateResiduals(sem_biomass_late2022, refit=FALSE))

sem_biomass_late2022_sqrt <- glmmTMB(sqrt(biomass) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_late, family = gaussian(), REML=F)
summary(sem_biomass_late2022_sqrt) #AIC=159.8
plot(simulateResiduals(sem_biomass_late2022_sqrt, refit=FALSE))

sem_biomass_late2022_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_late, family = gaussian(), REML=F)
summary(sem_biomass_late2022_log) #AIC=86.7
plot(simulateResiduals(sem_biomass_late2022_log, refit=FALSE))


AICc(sem_biomass_late2022, sem_biomass_late2022_sqrt, sem_biomass_late2022_log)

#best model
sem_biomass_late2022_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2022_late, family = gaussian(), REML=T)
summary(sem_biomass_late2022_log)
plot(simulateResiduals(sem_biomass_late2022_log, refit=FALSE))
```



vegetation height 2022 - early
```{r}
sem_veghgt_early2022 <- glmmTMB(veghgt_ave ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_early, family = gaussian(), REML=F)
summary(sem_veghgt_early2022) #AIC=400.1
plot(simulateResiduals(sem_veghgt_early2022, refit=FALSE))

sem_veghgt_early2022_sqrt <- glmmTMB(sqrt(veghgt_ave) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_early, family = gaussian(), REML=F)
summary(sem_veghgt_early2022_sqrt) #AIC=153.8
plot(simulateResiduals(sem_veghgt_early2022_sqrt, refit=FALSE))

sem_veghgt_early2022_log <- glmmTMB(log(veghgt_ave+1) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_early, family = gaussian(), REML=F)
summary(sem_veghgt_early2022_log) #AIC=81.8
plot(simulateResiduals(sem_veghgt_early2022_log, refit=FALSE))

AICc(sem_veghgt_early2022, sem_veghgt_early2022_sqrt, sem_veghgt_early2022_log)

#best model
sem_veghgt_early2022_sqrt <- glmmTMB(sqrt(veghgt_ave) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_early, family = gaussian(), REML=T)
summary(sem_veghgt_early2022_sqrt) #AIC=153.8
plot(simulateResiduals(sem_veghgt_early2022_sqrt, refit=FALSE))
```

vegetation height 2022 - mid
```{r}
sem_veghgt_mid2022 <- glmmTMB(veghgt_ave ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_mid, family = gaussian(), REML=FALSE)
summary(sem_veghgt_mid2022) #AIC=301.7
plot(simulateResiduals(sem_veghgt_mid2022, refit=FALSE))

sem_veghgt_mid2022_sqrt <- glmmTMB(sqrt(veghgt_ave) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_mid, family = gaussian(), REML=F)
summary(sem_veghgt_mid2022_sqrt) #AIC=95.2
plot(simulateResiduals(sem_veghgt_mid2022_sqrt, refit=FALSE))

sem_veghgt_mid2022_log <- glmmTMB(log(veghgt_ave+1) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_mid, family = gaussian(), REML=FALSE)
summary(sem_veghgt_mid2022_log) #AIC=27.2
plot(simulateResiduals(sem_veghgt_mid2022_log, refit=FALSE))

AICc(sem_veghgt_mid2022, sem_veghgt_mid2022_sqrt, sem_veghgt_mid2022_log)

#best model
sem_veghgt_mid2022_log <- glmmTMB(log(veghgt_ave+1) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_mid, family = gaussian(), REML=T)
summary(sem_veghgt_mid2022_log) 
plot(simulateResiduals(sem_veghgt_mid2022_log, refit=FALSE))
```

vegetation height 2022 - late
```{r}
sem_veghgt_late2022 <- glmmTMB(veghgt_ave ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_late, family = gaussian(), REML=F)
summary(sem_veghgt_late2022) #AIC=213.1
plot(simulateResiduals(sem_veghgt_late2022, refit=FALSE))

sem_veghgt_late2022_sqrt <- glmmTMB(sqrt(veghgt_ave) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_late, family = gaussian(), REML=F)
summary(sem_veghgt_late2022_sqrt) #AIC=65.2
plot(simulateResiduals(sem_veghgt_late2022_sqrt, refit=FALSE))

sem_veghgt_late2022_log <- glmmTMB(log(veghgt_ave+1) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_late, family = gaussian(), REML=F)
summary(sem_veghgt_late2022_log) #AIC=20.9
plot(simulateResiduals(sem_veghgt_late2022_log, refit=FALSE))

AICc(sem_veghgt_late2022, sem_veghgt_late2022_sqrt, sem_veghgt_late2022_log)

#best model
sem_veghgt_late2022_log <- glmmTMB(log(veghgt_ave+1) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2022_late, family = gaussian(), REML=T)
summary(sem_veghgt_late2022_log) 
plot(simulateResiduals(sem_veghgt_late2022_log, refit=FALSE))
```



GPI 2022 - early
```{r}
sem_gpi_early2022 <- glmmTMB(GPI200_ave ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_early, family = gaussian(), REML=FALSE)
summary(sem_gpi_early2022) #AIC=289.8
plot(simulateResiduals(sem_gpi_early2022, refit=FALSE))

sem_gpi_early2022_sqrt <- glmmTMB(sqrt(GPI200_ave) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_early, family = gaussian(), REML=FALSE)
summary(sem_gpi_early2022_sqrt) #AIC=-220.5
plot(simulateResiduals(sem_gpi_early2022_sqrt, refit=FALSE))

sem_gpi_early2022_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_early, family = gaussian(), REML=FALSE)
summary(sem_gpi_early2022_log) #AIC=-553.6
plot(simulateResiduals(sem_gpi_early2022_log, refit=FALSE))

AICc(sem_gpi_early2022, sem_gpi_early2022_sqrt, sem_gpi_early2022_log)

# best model
sem_gpi_early2022_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_early, family = gaussian(), REML=T)
summary(sem_gpi_early2022_log) 
plot(simulateResiduals(sem_gpi_early2022_log, refit=FALSE))
```

GPI 2022 - mid
```{r}
sem_gpi_mid2022 <- glmmTMB(GPI200_ave ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_mid, family = gaussian(), REML=FALSE)
summary(sem_gpi_mid2022) #AIC=217.2
plot(simulateResiduals(sem_gpi_mid2022, refit=FALSE))

sem_gpi_mid2022_sqrt <- glmmTMB(sqrt(GPI200_ave) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_mid, family = gaussian(), REML=FALSE)
summary(sem_gpi_mid2022_sqrt) #AIC=-181.5
plot(simulateResiduals(sem_gpi_mid2022_sqrt, refit=FALSE))

sem_gpi_mid2022_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_mid, family = gaussian(), REML=FALSE)
summary(sem_gpi_mid2022_log) #AIC=-441.7
plot(simulateResiduals(sem_gpi_mid2022_log, refit=FALSE))

AICc(sem_gpi_mid2022, sem_gpi_mid2022_sqrt, sem_gpi_mid2022_log)

# best model
sem_gpi_mid2022_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_mid, family = gaussian(), REML=T)
summary(sem_gpi_mid2022_log) 
plot(simulateResiduals(sem_gpi_mid2022_log, refit=FALSE))
```

GPI 2022 - late
```{r}
sem_gpi_late2022 <- glmmTMB(GPI200_ave ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_late, family = gaussian(), REML=FALSE)
summary(sem_gpi_late2022) #AIC=188.5
plot(simulateResiduals(sem_gpi_late2022, refit=FALSE))

sem_gpi_late2022_sqrt <- glmmTMB(sqrt(GPI200_ave) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_late, family = gaussian(), REML=FALSE)
summary(sem_gpi_late2022_sqrt) #AIC=-122.5
plot(simulateResiduals(sem_gpi_late2022_sqrt, refit=FALSE))

sem_gpi_late2022_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_late, family = gaussian(), REML=FALSE)
summary(sem_gpi_late2022_log) #AIC=-325.4
plot(simulateResiduals(sem_gpi_late2022_log, refit=FALSE))

AICc(sem_gpi_late2022, sem_gpi_late2022_sqrt, sem_gpi_late2022_log)

# best model
sem_gpi_late2022_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2022_late, family = gaussian(), REML=T)
summary(sem_gpi_late2022_log) 
plot(simulateResiduals(sem_gpi_late2022_log, refit=FALSE))
```


Soil moisture models do not look good

soil moisture 2022 - early
```{r}
sem_sm_early2022 <- glmmTMB(sm_ave ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_early, family = gaussian(), REML=FALSE)
summary(sem_sm_early2022) #AIC=531.4
plot(simulateResiduals(sem_sm_early2022, refit=FALSE))

sem_sm_early2022_sqrt <- glmmTMB(sqrt(sm_ave) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_early, family = gaussian(), REML=FALSE)
summary(sem_sm_early2022_sqrt) #AIC=212.5
plot(simulateResiduals(sem_sm_early2022_sqrt, refit=FALSE))

sem_sm_early2022_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_early, family = gaussian(), REML=FALSE)
summary(sem_sm_early2022_log) #AIC=76.3
plot(simulateResiduals(sem_sm_early2022_log, refit=FALSE))

AICc(sem_sm_early2022, sem_sm_early2022_sqrt, sem_sm_early2022_log)

# best model
sem_sm_early2022_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_early, family = gaussian(), REML=T)
summary(sem_sm_early2022_log) #AIC=76.3
plot(simulateResiduals(sem_sm_early2022_log, refit=FALSE))
```

soil moisture 2022 - mid
```{r}
sem_sm_mid2022 <- glmmTMB(sm_ave ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_mid, family = gaussian(), REML=FALSE)
summary(sem_sm_mid2022) #AIC=397.5
plot(simulateResiduals(sem_sm_mid2022, refit=FALSE))

sem_sm_mid2022_sqrt <- glmmTMB(sqrt(sm_ave) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_mid, family = gaussian(), REML=FALSE)
summary(sem_sm_mid2022_sqrt) #AIC=145.3
plot(simulateResiduals(sem_sm_mid2022_sqrt, refit=FALSE))

sem_sm_mid2022_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_mid, family = gaussian(), REML=FALSE)
summary(sem_sm_mid2022_log) #AIC=34.1
plot(simulateResiduals(sem_sm_mid2022_log, refit=FALSE))

AICc(sem_sm_mid2022, sem_sm_mid2022_sqrt, sem_sm_mid2022_log)

# best model
sem_sm_mid2022_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_mid, family = gaussian(), REML=T)
summary(sem_sm_mid2022_log)
plot(simulateResiduals(sem_sm_mid2022_log, refit=FALSE))
```

soil moisture 2022 - late
```{r}
sem_sm_late2022 <- glmmTMB(sm_ave ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_late, family = gaussian(), REML=FALSE)
summary(sem_sm_late2022) #AIC=231.7
plot(simulateResiduals(sem_sm_late2022, refit=FALSE))

sem_sm_late2022_sqrt <- glmmTMB(sqrt(sm_ave) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_late, family = gaussian(), REML=FALSE)
summary(sem_sm_late2022_sqrt) #AIC=77.2
plot(simulateResiduals(sem_sm_late2022_sqrt, refit=FALSE))

sem_sm_late2022_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_late, family = gaussian(), REML=FALSE)
summary(sem_sm_late2022_log) #AIC=27.6
plot(simulateResiduals(sem_sm_late2022_log, refit=FALSE))

AICc(sem_sm_late2022, sem_sm_late2022_sqrt, sem_sm_late2022_log)

# best model
sem_sm_late2022_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2022_late, family = gaussian(), REML=T)
summary(sem_sm_late2022_log)
plot(simulateResiduals(sem_sm_late2022_log, refit=FALSE))
```






Biomass 2023 - early 
```{r}
sem_biomass_early2023 <- glmmTMB(biomass ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_biomass_early2023) #AIC=605.6
plot(simulateResiduals(sem_biomass_early2023, refit=FALSE))

sem_biomass_early2023_sqrt <- glmmTMB(sqrt(biomass) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_biomass_early2023_sqrt) #AIC=266.0
plot(simulateResiduals(sem_biomass_early2023_sqrt, refit=FALSE))

sem_biomass_early2023_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_biomass_early2023_log) #AIC=182.4
plot(simulateResiduals(sem_biomass_early2023_log, refit=FALSE))


AICc(sem_biomass_early2023, sem_biomass_early2023_sqrt, sem_biomass_early2023_log)

#best model
sem_biomass_early2023_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_early, family = gaussian(), REML=T)
summary(sem_biomass_early2023_log) 
plot(simulateResiduals(sem_biomass_early2023_log, refit=FALSE))
```

biomass 2023 - mid
```{r}
sem_biomass_mid2023 <- glmmTMB(biomass ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_biomass_mid2023) #AIC=545.7
plot(simulateResiduals(sem_biomass_mid2023, refit=FALSE))

sem_biomass_mid2023_sqrt <- glmmTMB(sqrt(biomass) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_biomass_mid2023_sqrt) #AIC=228.6
plot(simulateResiduals(sem_biomass_mid2023_sqrt, refit=FALSE))

sem_biomass_mid2023_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_biomass_mid2023_log) #AIC=102.1
plot(simulateResiduals(sem_biomass_mid2023_log, refit=FALSE))


AICc(sem_biomass_mid2023, sem_biomass_mid2023_sqrt, sem_biomass_mid2023_log)

#best model
sem_biomass_mid2023_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_mid, family = gaussian(), REML=T)
summary(sem_biomass_mid2023_log) 
plot(simulateResiduals(sem_biomass_mid2023_log, refit=FALSE))
```

biomass 2023 - late
```{r}
sem_biomass_late2023 <- glmmTMB(biomass ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_biomass_late2023) #AIC=316.4
plot(simulateResiduals(sem_biomass_late2023, refit=FALSE))

sem_biomass_late2023_sqrt <- glmmTMB(sqrt(biomass) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_biomass_late2023_sqrt) #AIC=142.2
plot(simulateResiduals(sem_biomass_late2023_sqrt, refit=FALSE))

sem_biomass_late2023_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_biomass_late2023_log) #AIC=85.3
plot(simulateResiduals(sem_biomass_late2023_log, refit=FALSE))


AICc(sem_biomass_late2023, sem_biomass_late2023_sqrt, sem_biomass_late2023_log)

#best model
sem_biomass_late2023_log <- glmmTMB(log(biomass+1) ~ veghgt_ave.sc + sm_ave.sc + GPI200_ave.sc + LUI2016.sc + mowing.sc + (1|rainfall_cat) + (1|windspeed_cat) + knmitemp_ave.sc, data = df2_2023_late, family = gaussian(), REML=T)
summary(sem_biomass_late2023_log) 
plot(simulateResiduals(sem_biomass_late2023_log, refit=FALSE))
```



vegetation height 2023 - early
```{r}
sem_veghgt_early2023 <- glmmTMB(veghgt_ave ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_veghgt_early2023) #AIC=674.6
plot(simulateResiduals(sem_veghgt_early2023, refit=FALSE))

sem_veghgt_early2023_sqrt <- glmmTMB(sqrt(veghgt_ave) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_veghgt_early2023_sqrt) #AIC=291.1
plot(simulateResiduals(sem_veghgt_early2023_sqrt, refit=FALSE))

sem_veghgt_early2023_log <- glmmTMB(log(veghgt_ave+1) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_veghgt_early2023_log) #AIC=164.6
plot(simulateResiduals(sem_veghgt_early2023_log, refit=FALSE))

AICc(sem_veghgt_early2023, sem_veghgt_early2023_sqrt, sem_veghgt_early2023_log)

#best model
sem_veghgt_early2023_log <- glmmTMB(log(veghgt_ave+1) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_early, family = gaussian(), REML=T)
summary(sem_veghgt_early2023_log)
plot(simulateResiduals(sem_veghgt_early2023_log, refit=FALSE))
```

vegetation height 2023 - mid
```{r}
sem_veghgt_mid2023 <- glmmTMB(veghgt_ave ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_veghgt_mid2023) #AIC=510.4
plot(simulateResiduals(sem_veghgt_mid2023, refit=FALSE))

sem_veghgt_mid2023_sqrt <- glmmTMB(sqrt(veghgt_ave) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_veghgt_mid2023_sqrt) #AIC=211.9
plot(simulateResiduals(sem_veghgt_mid2023_sqrt, refit=FALSE))

sem_veghgt_mid2023_log <- glmmTMB(log(veghgt_ave+1) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_veghgt_mid2023_log) #AIC=97.9
plot(simulateResiduals(sem_veghgt_mid2023_log, refit=FALSE))

AICc(sem_veghgt_mid2023, sem_veghgt_mid2023_sqrt, sem_veghgt_mid2023_log)

#best model
sem_veghgt_mid2023_log <- glmmTMB(log(veghgt_ave+1) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_mid, family = gaussian(), REML=T)
summary(sem_veghgt_mid2023_log) 
plot(simulateResiduals(sem_veghgt_mid2023_log, refit=FALSE))
```

vegetation height 2023 - late
```{r}
sem_veghgt_late2023 <- glmmTMB(veghgt_ave ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_veghgt_late2023) #AIC=266.1
plot(simulateResiduals(sem_veghgt_late2023, refit=FALSE))

sem_veghgt_late2023_sqrt <- glmmTMB(sqrt(veghgt_ave) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_veghgt_late2023_sqrt) #AIC=98.8
plot(simulateResiduals(sem_veghgt_late2023_sqrt, refit=FALSE))

sem_veghgt_late2023_log <- glmmTMB(log(veghgt_ave+1) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_veghgt_late2023_log) #AIC=42.2
plot(simulateResiduals(sem_veghgt_late2023_log, refit=FALSE))

AICc(sem_veghgt_late2023, sem_veghgt_late2023_sqrt, sem_veghgt_late2023_log)

#best model
sem_veghgt_late2023_log <- glmmTMB(log(veghgt_ave+1) ~ sm_ave.sc + (1|rainfall_cat) + knmitemp_ave.sc + mowing.sc, data = df2_2023_late, family = gaussian(), REML=T)
summary(sem_veghgt_late2023_log)
plot(simulateResiduals(sem_veghgt_late2023_log, refit=FALSE))
```



GPI 2023 - early
```{r}
sem_gpi_early2023 <- glmmTMB(GPI200_ave ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_gpi_early2023) #AIC=395.5
plot(simulateResiduals(sem_gpi_early2023, refit=FALSE))

sem_gpi_early2023_sqrt <- glmmTMB(sqrt(GPI200_ave) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_gpi_early2023_sqrt) #AIC=-314.2
plot(simulateResiduals(sem_gpi_early2023_sqrt, refit=FALSE))

sem_gpi_early2023_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_gpi_early2023_log) #AIC=-777.5
plot(simulateResiduals(sem_gpi_early2023_log, refit=FALSE))

AICc(sem_gpi_early2023, sem_gpi_early2023_sqrt, sem_gpi_early2023_log)

# best model
sem_gpi_early2023_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_early, family = gaussian(), REML=T)
summary(sem_gpi_early2023_log) 
plot(simulateResiduals(sem_gpi_early2023_log, refit=FALSE))
```

GPI 2023 - mid
```{r}
sem_gpi_mid2023 <- glmmTMB(GPI200_ave ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_gpi_mid2023) #AIC=278.9
plot(simulateResiduals(sem_gpi_mid2023, refit=FALSE))

sem_gpi_mid2023_sqrt <- glmmTMB(sqrt(GPI200_ave) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_gpi_mid2023_sqrt) #AIC=-231.5
plot(simulateResiduals(sem_gpi_mid2023_sqrt, refit=FALSE))

sem_gpi_mid2023_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_gpi_mid2023_log) #AIC=-564.6
plot(simulateResiduals(sem_gpi_mid2023_log, refit=FALSE))

AICc(sem_gpi_mid2023, sem_gpi_mid2023_sqrt, sem_gpi_mid2023_log)

# best model
sem_gpi_mid2023_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_mid, family = gaussian(), REML=T)
summary(sem_gpi_mid2023_log) 
plot(simulateResiduals(sem_gpi_mid2023_log, refit=FALSE))
```

GPI 2023 - late
```{r}
sem_gpi_late2023 <- glmmTMB(GPI200_ave ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_gpi_late2023) #AIC=198.1
plot(simulateResiduals(sem_gpi_late2023, refit=FALSE))

sem_gpi_late2023_sqrt <- glmmTMB(sqrt(GPI200_ave) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_gpi_late2023_sqrt) #AIC=-128.9
plot(simulateResiduals(sem_gpi_late2023_sqrt, refit=FALSE))

sem_gpi_late2023_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_gpi_late2023_log) #AIC=-342.3
plot(simulateResiduals(sem_gpi_late2023_log, refit=FALSE))

AICc(sem_gpi_late2023, sem_gpi_late2023_sqrt, sem_gpi_late2023_log)

# best model
sem_gpi_late2023_log <- glmmTMB(log(GPI200_ave+1) ~ sm_ave.sc + veghgt_ave.sc, data = df2_2023_late, family = gaussian(), REML=T)
summary(sem_gpi_late2023_log)
plot(simulateResiduals(sem_gpi_late2023_log, refit=FALSE))
```



Soil moisture models do not look good

soil moisture 2023 - early
```{r}
sem_sm_early2023 <- glmmTMB(sm_ave ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_sm_early2023) #AIC=744.4
plot(simulateResiduals(sem_sm_early2023, refit=FALSE))

sem_sm_early2023_sqrt <- glmmTMB(sqrt(sm_ave) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_sm_early2023_sqrt) #AIC=280.1
plot(simulateResiduals(sem_sm_early2023_sqrt, refit=FALSE))

sem_sm_early2023_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_early, family = gaussian(), REML=FALSE)
summary(sem_sm_early2023_log) #AIC=67.2
plot(simulateResiduals(sem_sm_early2023_log, refit=FALSE))

AICc(sem_sm_early2023, sem_sm_early2023_sqrt, sem_sm_early2023_log)

# best model
sem_sm_early2023_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_early, family = gaussian(), REML=T)
summary(sem_sm_early2023_log) #AIC=76.3
plot(simulateResiduals(sem_sm_early2023_log, refit=FALSE))
```

soil moisture 2023 - mid
```{r}
sem_sm_mid2023 <- glmmTMB(sm_ave ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_sm_mid2023) #AIC=506.2
plot(simulateResiduals(sem_sm_mid2023, refit=FALSE))

sem_sm_mid2023_sqrt <- glmmTMB(sqrt(sm_ave) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_sm_mid2023_sqrt) #AIC=223.9
plot(simulateResiduals(sem_sm_mid2023_sqrt, refit=FALSE))

sem_sm_mid2023_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_mid, family = gaussian(), REML=FALSE)
summary(sem_sm_mid2023_log) #AIC=125.5
plot(simulateResiduals(sem_sm_mid2023_log, refit=FALSE))

AICc(sem_sm_mid2023, sem_sm_mid2023_sqrt, sem_sm_mid2023_log)

# best model
sem_sm_mid2023_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_mid, family = gaussian(), REML=T)
summary(sem_sm_mid2023_log)
plot(simulateResiduals(sem_sm_mid2023_log, refit=FALSE))
```

soil moisture 2023 - late
```{r}
sem_sm_late2023 <- glmmTMB(sm_ave ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_sm_late2023) #AIC=337.6
plot(simulateResiduals(sem_sm_late2023, refit=FALSE))


sem_sm_late2023_sqrt <- glmmTMB(sqrt(sm_ave) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_sm_late2023_sqrt) #AIC=155.1
plot(simulateResiduals(sem_sm_late2023_sqrt, refit=FALSE))

sem_sm_late2023_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_late, family = gaussian(), REML=FALSE)
summary(sem_sm_late2023_log) #AIC=95.8
plot(simulateResiduals(sem_sm_late2023_log, refit=FALSE))

AICc(sem_sm_late2023, sem_sm_late2023_sqrt, sem_sm_late2023_log)

# best model
sem_sm_late2023_log <- glmmTMB(log(sm_ave+1) ~ knmitemp_ave.sc + (1|rainfall_cat), data = df2_2023_late, family = gaussian(), REML=T)
summary(sem_sm_late2023_log)
plot(simulateResiduals(sem_sm_late2023_log, refit=FALSE))
```
